---
title: "Getting started with gevitR"
author: "Anamaria Crisan"
date: '2019-01-09'
output: html_document
---

## Inputting data using gevitR

gevitR provides a simple input function that allows users to easily input various types of data. By using the gevitR input functions its possible standardize how data is represented and it makes it easier to plot your data.

To load your own data, and get it ready for gevitR, you need to do the following:

```{r,eval=FALSE,message=FALSE,warning=FALSE}
input_data(file = "path/to/your/file.ext", dataType = "choose_data_type")

# where choose_data_type can be one of table, tree, dna, spatial, or image
```

To demonstrate the features of gevitR we are using data from the 2014-2016 Ebola Outbreak. This data is available from several sources:

Metadata and tree data is available via micoreact : 
https://microreact.org/project/west-african-ebola-epidemic

Genomic data (guinea only) is available via the following github repo: https://github.com/nickloman/ebov/tree/master/consensus/good

Spatial (shape file) data is available via : https://data.humdata.org

This data is automatically downloaded when gevitR is installed, and can be accessed using the commands below:

```{r,message=FALSE,warning=FALSE}
devtools::load_all() #temporary once things are done

#Table data
tab_dat<-input_data(file = system.file("extdata", "ebov_metadata.csv", package = "gevitR"),dataType = "table")

#Tree data
tree_dat<-input_data(file = system.file("extdata", "ebov_tree.nwk", package = "gevitR"),dataType = "tree")

#Genomic data
genomic_dat<-input_data(file = system.file("extdata", "ebov_GIN_genomic.fasta", package = "gevitR"),dataType = "dna")

#Shape files
#Shape files require that .shp,.shx,and .prj files at a minimun to be in the same directory
#to add metadata to the shape file, you can also add .dbf files
gin_shape_dat<-input_data(file = system.file("extdata", "gin_admbnda_adm1_ocha_itos.shp", package = "gevitR"),dataType = "spatial")
lbr_shape_dat<-input_data(file = system.file("extdata", "lbr_admbnda_adm1_ocha.shp", package = "gevitR"),dataType = "spatial")
sle_shape_dat<-input_data(file = system.file("extdata", "sle_admbnda_adm1_1m_gov_ocha_20161017.shp", package = "gevitR"),dataType = "spatial")
```

## Essential gevitR commands

**Some helper commands**
These commands let you know what you parameters you can provide each function to plot. To see examples of different charts type check out http://gevit.net
```{r,message=FALSE,warning=FALSE}

base_chart_types() #to see all of the supported chart types
base_chart_specs(chart_type="bar") #to see all of the required and options specifications

```
